Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-03-11 ms.tex
random stuff

A better understanding of what mechanisms regulate accretion onto
quiescent SMBHs would have implications for a variety of observations,
such as constraints on the occupation fraction of SMBHs in low mass
galaxies.  \citet{Miller+15} use the average relationship between the
nuclear X-ray luminosities, $L_{X}$, of a sample of elliptical
galaxies and the black hole mass to infer tentative evidence for a
SMBH occupation fraction less than unity in galaxies with stellar
masses $M_{\star} \lesssim 10^{10}M_{\odot}$ ($M_{\bullet} \lesssim
10^{7}M_{\odot}$).  This method relies on extrapolating a power-law
fit of the $L_X-\Mbh$ distribution to low values of $L_{\rm X}$ below
the detection threshold, a potentially problematic assumption if
different physical processes control the accretion rates onto the
lowest mass SMBHs.

@@

\citet{Ho:2009a} estimates the SMBH accretion rates in a sample of
early-type galaxies using X-ray observations to determine the
Bondi accretion rate and based on estimates for the mass loss rate of
evolved stars.  Both methods lead to the conclusion that the available
gas reservoir is more than sufficient to power the observed
low-luminosity AGN, assuming the standard $\sim$ 10 per cent radiative
efficiency for thin disk accretion.  Much
evidence now suggest that low-luminosity AGN result from accretion
proceeding in a radiatively inefficient mode
(\citealt{Yuan&Narayan14}), due either to the advection of
gravitationally-released energy across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to the low efficiency with which the inflowing gas ultimately reaching the viscinity of the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

@@

Past studies, employing multi-dimensional numerical hydrodynamics and
including variety of (parameterized) physical effects, have focused on
massive elliptical galaxies (e.g.~\citealt{Ciotti&Ostriker07};
\citealt{Ciotti+10}).  A key finding of these works is the periodic
development of cooling instabilities on galactic scales, which
temporarily increase the gas accretion rate towards the nucleus until
feedback becomes sufficiently effective to shut off the flow and halt
SMBH growth.  Here we instead focus on time-independent models,
neglecting radiative cooling.  This approach allows us to
systematically explore the relevant parameter space and to derive
simple analytic expressions that prove useful in exploring whether
cooling instabilties manifest across the expected range of galaxy
properties, or whether other (non-AGN) forms of feedback can result in
steady accretion.  Even if cooling instabilties develop on galactic
scales over longer timescales, e.g. $\sim$ Gyr, a state of
quasi-steady accretion may exist between these inflow events on
smaller radial scales comparable to the sphere of influence.  We are
thus motivated to explore whether steady flow is achieved within the
nucleus under any circumstances, or whether the gas content is always
building towards instability.
%A chief goal of our analysis is to provide a framework for interpreting the properties of quiescent galactic nuclei given the comparatively well understood physics of stellar winds and supernova heating.

@@

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our numerical results.  In $\S\ref{sec:discussion}$ we discuss the implications of our results, to topics including...   In $\S\ref{sec:conclusions}$ we summarize our conclusions.  Table \ref{table:definitions} provides the definitions of commonly used variables.  Appendix \ref{app:rs} provides useful analytic results, while Appendix \ref{app:windheat} describes our method for calculating stellar wind heating and mass input in greater detail.

@@

Radiative cooling contributes an additional term to equation (\ref{eq:dsdt}) of the form
\begin{equation}
\dot{q}_{\rm rad} = -\Lambda(T)n^{2},
\label{eq:qdot_rad}
\end{equation}
where $n = \rho/\mu m_p$ is the gas number density and $\Lambda(T)$ is the cooling function.  We neglect radiative cooling in our baseline calculations, despite the fact that this is not justified when the wind heating $\vwO$ is low or if the mass return rate $\eta$ is high.  However, once radiative cooling becomes important compared to other sources of heating and cooling, its presence leads to thermal instability (e.g.~\citealt{McCourt+12}; $\S\ref{sec:instability}$) that cannot be accounted for by our 1D
time-independent model.  A main goal of this work is to use solutions neglecting radiation to ascertain over what range of physical conditions cooling instabilities are most likely to develop ($\S\ref{sec:instability}$).

@@

Continuity of the entropy derivative at the stagnation radius where $v = 0$ requires that the temperature at this location be given by (eq.~[\ref{eq:first_law}])
\begin{align}
T|_{r_{s}}& = \frac{\gamma-1}{\gamma}\frac{\mu m_p
  \tilde{v}_{w}^{2}}{2k} \nonumber\\
 &\approx \frac{2}{5} \frac{7-2\nu}{7} \frac{\mu m_p
   v_{w}^{2}}{2k}\nonumber\\ 
 &\approx
 \begin{cases}
  4.5\times 10^6\ \, v_{500}^2 \,\,{\rm K} & \text{core} \\
  5.3\times 10^6\ \, v_{500}^2 \,\,{\rm K} & \text{cusp},
 \end{cases}
\label{eq:Tanalytic}
\end{align}
where we have assumed $\gamma=5/3$, $v_{500} \equiv v_{w}/(500$ km
s$^{-1}$) and $\densSlope \equiv -d{\rm ln}\rho/d{\rm ln\,r}|_{r_{\rm
    s}}$ is the density power-law slope at $r = r_{\rm s}$.
Empirically, we find that
\begin{align}
\densSlope \simeq \frac{1}{6} \left(4 \Gamma+3\right).
\label{eq:densSlope}
\end{align}

@@

In the opposite limit of weak heating ($v_{w} \ll \sigma$), the
stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not
exist at all, formally implying that the entire ISM of the galaxy is
accreting.  This result makes intuitive sense: gas is supplied to the
nuclear by stars which are gravitationally bound to the black hole, so
outflows are possible only if the specific heating rate $\sim
v_{w}^{2}$ significantly exceeds the specific gravitational binding
energy $\sim \sigma^{2}$.

@@

The gas density at the stagnation radius $\rho|_{r_{\rm s}}$ is more challenging to estimate accurately.  Using an alternative estimate of $\dot{M}$ as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}|_{r_{\rm s}} = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$,
\begin{align}
  &\dot{M}\sim\frac{(4 \pi/3) \rs^3 \rho|_{r_{\rm s}}}{\tff|_{r_{\rm s}}},
  \label{eq:mdot_gas}
\end{align}
 in conjuction with equations (\ref{eq:mdot_analytic}) and (\ref{eq:rs_simple}), one derives
\begin{align}
  \rho|_{r_{\rm s}}\approx
  \begin{cases}
    6 \times 10^{-26} \Mbheight^{-0.2} v_{500}^{-0.8}  \eta_{0.02} \,
    \, {\rm g \, cm^{-3}}& \text{core},\\
    10 \times 10^{-26}  \Mbheight^{-0.5} v_{500}^{0.6}  \eta_{0.02} \,\, {\rm g \,cm^{-3}} & \text{cusp}.
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate (eq.~[\ref{eq:qdotcond}]) to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}}
\tilde{v}_{w}^{2}}
 \left(\frac{1}{1+\frac{\kappa_0 T^{7/2}|_{r_{\rm s}}}{5 \rs
    \phi \rho c_s^3}}\right)
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  2 \phi & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}),
(\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim
1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the
stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    6.1 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker
break radius and for clarity we have not distinguished between cusp
and core galaxies.  Equation (\ref{eq:conduction}) shows that even in
the case of saturated conduction, the neglect of heat conduction
results in at most an order unity correction for expected {\bf (AG:
  this sounds a bit too sharp perhaps replace with the word
  plausible. Perhaps we should also mention the phi=1 solutions have
  acausal conductivity (perhaps they are unphysical?).} values of
the saturation parameter $\phi < 0.1$, justifying its neglect in our
analysis.  Numerical experiments including conductivity confirm this
($\S\ref{sec:numerical}$).

@@

Radiative cooling can, however, be stabilized if it is overwhelmed by
other sources of cooling, namely the term $\propto -q c_{s}^{2}
\propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]).
Neglecting radiative cooling, this term is balanced at the stagnation
radius by the wind heating term, $\dot{q}_{\rm heat} = q
\tilde{v}_{w}^{2}/2$.  Thermal stability near the stagnation radius
can thus be assessed by comparing the ratio of wind heating to
radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   380 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}

where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs})
for the gas and stellar densities at the stagnation radius,
respectively.  We have approximated the cooling function for $T <
2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$ assuming solar
metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).

@@

Thermal stability ($\dot{q}_{\rm heat} > |\dot{q}_{\rm rad}|$; $t_{\rm
  cool} > t_{\rm ff}$) near the stagnation radius thus requires a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
\sigma,     
  \end{cases}
  \label{eq:cooling3}
\end{align}
The requirement $v_{\rm w} \gtrsim \sigma$ results because the top two lines are derived using equation (\ref{eq:rs_simple}) for $r_s$.  However, when $v_{\rm w} \lesssim \sigma$ the stagnation radius moves to very large radii, or may not exist at all (see also Fig.~\ref{fig:stag}), resulting in much higher gas densities and inevitable cooling instability. 
%Our numerical results in $\S\ref{sec:numerical}$ demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm CI} \sim 200-300$ km s$^{-1}$ is generally sufficient for stability at all radii (Fig.~\ref{fig:cooling}).

@@

  \citet{EmsellemCappellari+:2007a} use two-dimensional kinematic data
  to measure the ratio of ordered to random motion in a sample of
  early type galaxies, which they quantify at each galactic radius $R$
  by the parameter
  \begin{align}
    \lambda_R \equiv \frac{\langle r|V|\rangle}{\langle R\sqrt{V^2+\sigma^2}\rangle} \underset{R \ll r_{\rm soi}}\sim \frac{V_{\phi}}{\sigma},
  \end{align}

  where $\sigma$ is the velocity dispersion and the brackets indicate
  a luminosity-weighted average.  The circularization radius of the
  accretion flow can be written in terms of $\lambda_R$ as

\begin{equation}
\frac{r_{\rm circ}}{r_{s}} \approx \frac{r_{s}}{r_{\rm soi}}\lambda_{R}^{2} \lesssim \lambda_{R}^{2},
\label{eq:rcirc}
\end{equation}

where we have used the definition $r_{\rm soi} \equiv
GM_{\bullet}/\sigma^{2}$ and in the second inequality have assumed
that $r_{\rm s} \lesssim r_{\rm soi}$, a condition which is satisfied
for the thermally-stable solutions of interest.

\citet{EmsellemCappellari+:2007a} (their Fig.~2) find that $\lambda_R$
is generally $< 0.1$ on radial scales $<$ 10 per cent of the galaxy
half-light radius and that $\lambda_R$ is with $R$ interior to this
point.  From equation (\ref{eq:rcirc}) we thus conclude that $r_{\rm
  circ} \lesssim 0.01r_{s}$, i.e.~that gas will circularize well
inside the inner sonic point of the flow (typically located at $r\lsim
0.1r_{\rm s}$; Fig.~\ref{fig:profiles}).  Angular momentum should only
become dynamically important for thermally stable flows in a region
that is causally disconnected from the outer part of the flow where
the mass accretion rate is determined.

>>>>

A better understanding of what mechanisms regulate accretion onto
quiescent SMBHs would have implications for a variety of observations,
such as constraints on the occupation fraction of SMBHs in low mass
galaxies.  \citet{Miller+15} use the average relationship between the
nuclear X-ray luminosities, $L_{X}$, of a sample of elliptical
galaxies and their associated SMBH masses to infer tentative evidence for a
SMBH occupation fraction less than unity in galaxies with stellar
masses $M_{\star} \lesssim 10^{10}M_{\odot}$ ($M_{\bullet} \lesssim
10^{7}M_{\odot}$).  This method relies on extrapolating a power-law
fit of the $L_X-\Mbh$ distribution to low values of $L_{\rm X}$ below
the detection threshold, a potentially problematic assumption if
different physical processes control the accretion rates onto the
lowest mass SMBHs.

@@

\citet{Ho:2009a} determines the SMBH accretion rates in a sample of
early-type galaxies using X-ray observations to determine the
Bondi accretion rate and based on estimates for the mass loss rate of
evolved stars.  Both methods lead to the conclusion that the available
gas reservoir is more than sufficient to power the observed
low-luminosity AGN, assuming the standard $\sim$ 10 per cent radiative
efficiency for thin disk accretion.  Much
evidence now suggest that low-luminosity AGN result from accretion
proceeding in a radiatively inefficient mode
(\citealt{Yuan&Narayan14}), due either to the advection of
gravitationally-released energy across the SMBH horizon
(e.g.~\citealt{Narayan&Yi95}) or due to the low efficiency with which the inflowing gas ultimately reaching the viscinity of the SMBH
(e.g.~\citealt{Blandford&Begelman99}; \citealt{Li+13}).

@@

Past studies, employing multi-dimensional numerical hydrodynamics and
including variety of (parameterized) physical effects, have focused on
massive elliptical galaxies (e.g.~\citealt{Ciotti&Ostriker07};
\citealt{Ciotti+10}).  A key finding of these works is the periodic
development of cooling instabilities on galactic scales, which
temporarily increase the gas accretion rate towards the nucleus until
feedback becomes effective to shut off the flow and halt
SMBH growth.  Here we instead focus on time-independent models,
neglecting radiative cooling.  This approach allows us to
systematically explore the relevant parameter space and to derive
simple analytic expressions that prove useful in exploring whether
cooling instabilties manifest across the expected range of galaxy
properties, or whether other (non-AGN) forms of feedback can result in
steady accretion.  Even if cooling instabilties develop on galactic
scales over longer $\sim$ Gyr timescales, a state of
quasi-steady accretion may exist between these inflow events on
smaller radial scales comparable to the sphere of influence.  We are
thus motivated to explore whether steady flow is achieved within the
nucleus under any circumstances, or whether the gas content is always
building towards instability.
%A chief goal of our analysis is to provide a framework for interpreting the properties of quiescent galactic nuclei given the comparatively well understood physics of stellar winds and supernova heating.

@@

This paper is organized as followed.  In $\S\ref{sec:model}$ we
describe our model, including the sample of galaxies used in our
analysis ($\S\ref{sec:gal_model}$) and our numerical procedure for
calculating the steady-state hydrodynamic profile of the CNM
($\S\ref{sec:hydro}$).  In $\S\ref{sec:results}$ we describe some
analytic results, while in $\S\ref{sec:numerical}$ we describe our numerical results.  In $\S\ref{sec:discussion}$ we discuss the implications of our results, to topics including...   In $\S\ref{sec:conclusions}$ we summarize our conclusions.  Table \ref{table:definitions} provides the definitions of commonly used variables.  Appendices \ref{app:rs} and XXX provides useful analytic results, while Appendix \ref{app:windheat} describes our method for calculating stellar wind heating and mass input in greater detail.

@@

Radiative cooling contributes an additional term to equation (\ref{eq:dsdt}) of the form
\begin{equation}
\dot{q}_{\rm rad} = -\Lambda(T)n^{2},
\label{eq:qdot_rad}
\end{equation}
where $n = \rho/\mu m_p$ is the gas number density and $\Lambda(T)$ is the cooling function.  We neglect radiative cooling in our baseline calculations, despite the fact that this is not justified when the wind heating $\vwO$ is low or if the mass return rate $\eta$ is high.  However, once radiative cooling becomes important compared to other sources of heating and cooling, its presence leads to thermal instability (e.g.~\citealt{McCourt+12}) that cannot be accounted for by our 1D
time-independent model.  A main goal of this work is to use solutions neglecting radiation to ascertain over what range of physical conditions cooling instabilities are most likely to develop ($\S\ref{sec:instability}$).

@@

Continuity of the entropy derivative at the stagnation radius where $v = 0$ requires that the temperature at this location be given by (eq.~[\ref{eq:first_law}])
\begin{align}
T|_{r_{s}}& = \frac{\gamma-1}{\gamma}\frac{\mu m_p
  \tilde{v}_{w}^{2}}{2k} \approx \frac{2}{5} \frac{7-2\nu}{7} \frac{\mu m_p
   v_{w}^{2}}{2k}\nonumber\\ 
 &\approx
 \begin{cases}
  4.5\times 10^6\ \, v_{500}^2 \,\,{\rm K} & \text{core} \\
  5.3\times 10^6\ \, v_{500}^2 \,\,{\rm K} & \text{cusp},
 \end{cases}
\label{eq:Tanalytic}
\end{align}
where we have assumed $\gamma=5/3$, $v_{500} \equiv v_{w}/(500$ km
s$^{-1}$) and $\densSlope \equiv -d{\rm ln}\rho/d{\rm ln\,r}|_{r_{\rm
    s}}$ is the density power-law slope at $r = r_{\rm s}$.
Empirically, we find that
\begin{align}
\densSlope \simeq \frac{1}{6} \left(4 \Gamma+3\right).
\label{eq:densSlope}
\end{align}

@@

In the opposite limit of weak heating ($v_{w} \ll \sigma$), the
stagnation radius moves to large radii $\gg r_{\rm soi}$ or may not
exist at all, formally implying that the entire ISM of the galaxy is
accreting.  This result makes intuitive sense: gas is supplied to the
nuclear by stars which are gravitationally bound to the black hole, so
outflows are possible only if the specific heating rate $\sim
v_{w}^{2}$ significantly exceeds the specific gravitational binding
energy $\sim \sigma^{2}$.  In fact, for core galaxies the specific heating rate necessary to produce an unbound outflow can greatly exceed $\sigma^{2}$ due to the additional gravitational potential of stars outside the sphere of influence (Appendix XXX).

@@

The gas density at the stagnation radius $\rho|_{r_{\rm s}}$ is more challenging to estimate accurately.  Using an alternative estimate of $\dot{M}$ as the gaseous mass within the stagnation radius divided by the free-fall time $t_{\rm ff}|_{r_{\rm s}} = (r_{\rm s}^{3}/GM_{\bullet})^{1/2}$,
\begin{align}
  &\dot{M}\sim\frac{(4 \pi/3) \rs^3 \rho|_{r_{\rm s}}}{\tff|_{r_{\rm s}}},
  \label{eq:mdot_gas}
\end{align}
 in conjuction with equations (\ref{eq:mdot_analytic}) and (\ref{eq:rs_simple}), one derives
\begin{align}
  \rho|_{r_{\rm s}}\approx
  \begin{cases}
    6 \times 10^{-26} \Mbheight^{-0.2} v_{500}^{-0.8}  \eta_{0.02} \,
    \, {\rm g \, cm^{-3}}& \text{core},\\
    1.0 \times 10^{-25}  \Mbheight^{-0.5} v_{500}^{0.6}  \eta_{0.02} \,\, {\rm g \,cm^{-3}} & \text{cusp}.
  \end{cases}
  \label{eq:rhors}
\end{align}

@@

Our analytic derviations neglect the effects of heat conduction, an assumption we now check.  The ratio of the magnitude of the conductive heating rate (eq.~[\ref{eq:qdotcond}]) to the external heating rate at the stagnation radius is given by
\begin{align}
  \left.\frac{\nabla\cdot(\kappa \nabla T)}{q v_{\rm
w}^{2}/2}\right|_{r_{\rm s}} &\sim \frac{2t_{\rm h}\kappa_0
T|_{r_{\rm s}}^{7/2}}{r_{\rm s}^{2}\eta \rho_{\star}|_{r_{\rm s}}
\tilde{v}_{w}^{2}}
 \left(\frac{1}{1+\frac{\kappa_0 T^{7/2}|_{r_{\rm s}}}{5 \rs
    \phi \rho c_s^3}}\right)
\nonumber \\ &\sim {\rm min}
  \begin{cases}
  20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.8} v_{500}^{6.8} &  \text{unsaturated (core)}\\
 20 \eta_{0.02}^{-1}
M_{\bullet,8}^{-0.5} v_{500}^{5.4} &  \text{unsaturated (cusp)}\\
  2 \phi & \text{saturated},
  \end{cases}
 \label{eq:conduction}
\end{align}
where the second equality makes use of equations (\ref{eq:Tanalytic}),
(\ref{eq:stag_simple}), and we have approximated $\nabla^{2} \sim
1/r_{\rm s}^{2}$, $\nabla \sim 1/r_{\rm s}$.  We have approximated the
stellar density profile as (eq.~[\ref{eq:rhostar}])
\begin{eqnarray}
  \rho_{\star}|_{r_{\rm s}} &\simeq& \frac{M_{\bullet}(2-\Gamma)}{4\pi r_{\rm soi}^{3}}\left(\frac{r_{\rm s}}{r_{\rm soi}}\right)^{-1-\Gamma} \nonumber \\
 &\approx& \begin{cases}
    6.1 \times 10^{-19}M_{\bullet,8}^{-1.2}v_{500}^{2.2}\,{\rm g\,cm^{-3}}
    & \text{core} \\
    2.1 \times 10^{-18} M_{\bullet,8}^{-1.5}v_{500}^{3.6}
    \,{\rm g\,cm^{-3}}  & \text{cusp}, 
  \end{cases}
  \label{eq:rhostarrs}
\end{eqnarray}
where the stagnation radius is assume to reside well inside the Nuker
break radius and for clarity we have not distinguished between cusp
and core galaxies.  Equation (\ref{eq:conduction}) shows that even in
the case of saturated conduction, the neglect of heat conduction
results in at most an order unity correction for plausible, causal values of
the saturation parameter $\phi < 0.1$, justifying its neglect in our
analysis.  Numerical experiments including conductivity confirm this
($\S\ref{sec:numerical}$).

@@

Radiative cooling can, however, be stabilized if it is overwhelmed by
other sources of cooling, namely the term $\propto -q c_{s}^{2}
\propto -qT$ in the entropy equation (eq.~[\ref{eq:dsdt}]).
Neglecting radiative cooling, this term is balanced at the stagnation
radius by the wind heating term, $\dot{q}_{\rm heat} = q
\tilde{v}_{w}^{2}/2$.  Thermal stability near the stagnation radius
can thus be assessed by comparing the ratio of wind heating to
radiative cooling $|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:qdot_rad}]),
\begin{align}
\left.\frac{\dot{q}_{\rm heat}}{|\dot{q}_{\rm rad}|}\right|_{\rm r_{\rm s}} \simeq
  \begin{cases}
   380 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.76}v_{500}^{7.2}  &, \text{core}\\
   590 \eta_{0.02}^{-1} M_{\bullet,8}^{-0.48}v_{500}^{5.8}  &, \text{cusp},     
  \end{cases}
  \label{eq:cooling2}
\end{align}
where we have used equations (\ref{eq:rhors}) and (\ref{eq:rhostarrs})
for the gas and stellar densities at the stagnation radius,
respectively.  We have approximated the cooling function for $T <
2\times 10^{7}$ K as $\Lambda(T) = 1.1 \times 10^{-22} \left(T/10^6
  \text{K}\right)^{-0.7} $erg cm$^3 $s$^{-1}$ assuming solar
metallicity gas (\citealt{Draine:2011a}; his Fig. 34.1).

@@

Thermal stability ($\dot{q}_{\rm heat} > |\dot{q}_{\rm rad}|$; $t_{\rm
  cool} > t_{\rm ff}$) near the stagnation radius thus requires a
critical minimum heating rate
\begin{align}
v_{w} > v_{\rm TI} \simeq
  {\rm max}\begin{cases}
   200 \eta_{0.02}^{0.14} M_{\bullet,8}^{0.11}\,{\rm km\,s^{-1}}  &, \text{core}\\
   200 \eta_{0.02}^{0.17} M_{\bullet,8}^{0.08}\,{\rm km\,s^{-1}}   &, \text{cusp}\\
f(r_b/r_{\rm soi})\sigma,     
  \end{cases}
  \label{eq:cooling3}
\end{align}
The requirement $v_{\rm w} \gtrsim \sigma$ results because the top two lines are derived using equation (\ref{eq:rs_simple}) for $r_s$.  However, when $v_{\rm w} \lesssim f(r_b/r_{\rm soi})\sigma$ the stagnation radius moves to very large radii, or may not exist at all (see also Fig.~\ref{fig:stag}), resulting in much higher gas densities and inevitable cooling instability.  The dimensionless function f is approximately unity for cusp galaxies, but increases $\propto (r_b/r_{\rm soi})^{1/2}$ for core galaxies (Fig.~XXX).  
%Our numerical results in $\S\ref{sec:numerical}$ demonstrate that $\tilde{v}_{w} \gtrsim v_{\rm CI} \sim 200-300$ km s$^{-1}$ is generally sufficient for stability at all radii (Fig.~\ref{fig:cooling}).

@@

  \citet{EmsellemCappellari+:2007a} use two-dimensional kinematic data
  to measure the ratio of ordered to random motion in a sample of
  early type galaxies, which they quantify at each galactic radius $R$
  by the parameter
  \begin{align}
    \lambda_R \equiv \frac{\langle r|V|\rangle}{\langle R\sqrt{V^2+\sigma^2}\rangle} \underset{R \ll r_{\rm soi}}\sim \frac{V_{\phi}}{\sigma},
  \end{align}
  where $\sigma$ is the velocity dispersion and the brackets indicate
  a luminosity-weighted average.  The circularization radius of the
  accretion flow can be written in terms of $\lambda_R$ as

\begin{equation}
\frac{r_{\rm circ}}{r_{s}} \approx \frac{r_{s}}{r_{\rm soi}}\lambda_{R}^{2} \lesssim \lambda_{R}^{2},
\label{eq:rcirc}
\end{equation}
where we have used the definition $r_{\rm soi} \equiv
GM_{\bullet}/\sigma^{2}$ and in the second inequality have assumed
that $r_{\rm s} \lesssim r_{\rm soi}$, a condition which is satisfied
for the thermally-stable solutions of interest.

\citet{EmsellemCappellari+:2007a} (their Fig.~2) find that $\lambda_R$
is generally $< 0.1$ on radial scales $<$ 10 per cent of the galaxy
half-light radius and that $\lambda_R$ decreases with decreasing $R$ interior to this point.  From equation (\ref{eq:rcirc}) we thus conclude that $r_{\rm
  circ} \lesssim 0.01r_{s}$, i.e.~that gas will circularize well
inside the inner sonic point of the flow (typically located at $r\lsim
0.1r_{\rm s}$; Fig.~\ref{fig:profiles}).  Angular momentum should only
become dynamically important for thermally stable flows in a region
that is causally disconnected from the outer part of the flow where
the mass accretion rate is determined.
